RNAseq aCRY pCRY data dive (separate)

### 2024-06-14 2_A_dive_aCRY_pCRY (separately)/2_A_dive_aCRY_pCRY_sep.Rmd
advanced volcanos

@@

@@

```{r f_topgenes}
topgenes_f <- function(res,p=0.5,bM=20,l2FC=1){
a <- subset(res, padj < p & baseMean > bM & abs(log2FoldChange) > l2FC)
a <- a[order(a$baseMean, decreasing = T),]
  a$rank.bm <- seq(1:length(rownames(a)))
a <- a[order(a$padj, decreasing = F),]
  a$rank.padj <- seq(1:length(rownames(a)))
a <- a[order(abs(a$log2FoldChange), decreasing = T),]
  a$rank.l2FC <- seq(1:length(rownames(a)))
a$rank.sum <- a$rank.l2FC+a$rank.bm+a$rank.padj
  a <- a[order(a$rank.sum),]
a
}
```

@@

Vulcano_SK <- function(res,
                       n=NULL,
                       list1=NULL,
                       list2=NULL,
                       list1.col=NULL,
                       list2.col=NULL,
                       list1.n=NULL,
                       list2.n=NULL,
                       xlim=NULL,
                       ylim=NULL,
                       ...)
{
# Name
n <- names(res)
# unlist results
if(is.list(res)){
  res <- res[[1]]
}
# define list 1
if(is.null(list1) ) {
  list1 <- topgenes_f(res,p=0.05,bM = 10,l2FC = 1)
}
# define list 2
if(is.null(list2) ) {
  list2 <- topgenes_f(res,p=0.005,bM = 100,l2FC = 2)
}
# list 1 color
if(is.null(list1.col) ) {
  list1.col <- "royalblue1"
}
# list 2 color
if(is.null(list2.col) ) {
  list2.col <- "royalblue4"
}
# name list 1
if(is.null(list1.n) ) {
  list1.n <- "list1"
}
# name list 2
if(is.null(list2.n) ) {
  list2.n <- "list2"
}

@@

# add symbol
res$symbol <- mcols(dds_list[[i]])$symbol

# define list groups
res$isl1 <- rownames(res) %in% rownames(list1)
res$isl2 <- rownames(res) %in% rownames(list2)

# order results (list 1 on top)
res <- res[order(abs(res$log2FoldChange), decreasing = T),]
res <- res[order(res$isl1, decreasing = F),]
res <- res[order(res$isl2, decreasing = F),]

@@

# name colors
keyvals <- ifelse(
    res$isl2 == TRUE, list2.col,
    ifelse(
    res$isl1 == TRUE, list1.col,
    'grey') )

keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == list1.col] <- list1.n
  names(keyvals)[keyvals == list2.col] <- list2.n
  names(keyvals)[keyvals == 'grey'] <- 'other'

ev_f <- EnhancedVolcano(res,
    x = 'log2FoldChange',
    y = 'padj',
    lab = res$symbol,
    selectLab  = res$symbol[which(names(keyvals) == list2.n)],
    labSize = 3,
    drawConnectors = TRUE,
    boxedLabels = FALSE,
    widthConnectors = 0.5,
    colConnectors = lcol,
    pointSize = 4.0,
    max.overlaps = 100,
    colCustom = keyvals,
    xlim = c(-xlim,xlim),
    ylim = c(0, ylim),
    ylab = "Padj (-log10)",
    title = n,
    subtitle = paste("list1:",nrow(list1),", list2: ",nrow(list2)),
    # sub = "SVF",
    # pCutoff = 10^(-60),
    # FCcutoff = 1,
    shapeCustom =shape,
    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
    legendLabels=c('Not sig.','|L2F| > 2','p-adj < 0.05',
                    'p-adj & L2F'),
    legendPosition = 'bottom',
    col=c('grey', pcol, pcol, pcol)
    )
ev_f

@@

# single gene
    d <- plotCounts(dds, gene=ROC40, intgroup=c("condition","experiment","genotype","treatment","clientName"), main=s,returnData=TRUE)

gcounts <- ggplot(d, aes(x = treatment, y = count, fill=condition, color=condition)) +
    geom_boxplot(color="black") +
    geom_point(shape=21,color="black",aes(fill=condition),position=position_dodge(width=0.75), alpha=1) +
  geom_text_repel(aes(label=clientName)) +
    scale_fill_manual(values=group.colors) +
    scale_color_manual(values=group.colors) +
    scale_y_continuous(trans = "log2") +
    labs(title = paste("ROC40"," in Δ",colData(dds)$experiment[1],sep = ""))
gcounts %>% print()

@@

# ggsave(paste(figures,"2023_12_counts_phots_aio_log2.pdf",sep="/"), plot = gcounts,
#   width = 12,
#   height = 8)

@@

## Make results
```{r, eval=FALSE, include=FALSE, echo=FALSE}

@@

top_genes_l <- lapply(res_list,topgenes_f,p=0.01, bM=100, l2FC=2) %>%  lapply(.,rownames)

@@

## Volcanos
### dark
```{r, eval=TRUE}

i <- 2

Vulcano_SK(res_ashr_list[[i]]["pcry_BL.vs.WT_BL"],xlim=5,ylim=50)

>>>>

library(VennDiagram)

@@

pub_figures <- paste(gitdir,"pub_figures",sep="/")

@@

```{r f_topgenes}

topgenes_f <- function(res,p=0.005,bM=10,l2FC=1){
a <- subset(res, padj < p & baseMean > bM & abs(log2FoldChange) > l2FC)
if(nrow(a)>0) {
a <- a[order(a$baseMean, decreasing = T),]
  a$rank.bm <- seq(1:length(rownames(a)))
a <- a[order(a$padj, decreasing = F),]
  a$rank.padj <- seq(1:length(rownames(a)))
a <- a[order(abs(a$log2FoldChange), decreasing = T),]
  a$rank.l2FC <- seq(1:length(rownames(a)))
a$rank.sum <- a$rank.l2FC+a$rank.bm+a$rank.padj
  a <- a[order(a$rank.sum),]
  }
a
}
```

@@

# res <- res_ashr_list[[i]][6]

Vulcano_SK <- function(res,
                       n=NULL,
                       list1=NULL,
                       list2=NULL,
                       list1.col=NULL,
                       list2.col=NULL,
                       ol.col=NULL,
                       list1.n=NULL,
                       list2.n=NULL,
                       xlim=NULL,
                       ylim=NULL,

                       ...)
{
# Name
if(is.null(n))  {
  if(is.list(res)){
      n <- names(res)
  } else {
    n <- "results"
  }
  }

# unlist results
if(is.list(res)){
  res <- res[[1]]
}

# define list 1
if(is.null(list1) ) {
  list1 <- topgenes_f(res=res,p=0.05,bM = 10,l2FC = 1) %>% rownames()
}
# define list 2
if(is.null(list2) ) {
  list2 <- topgenes_f(res=res, p=0.01, l2FC = 2, bM=10) %>% rownames()
}
# list 1 color
if(is.null(list1.col) ) {
  list1.col <- "royalblue1"
}
# list 2 color
if(is.null(list2.col) ) {
  list2.col <- "royalblue4"
}
# list ol color
if(is.null(ol.col) ) {
  ol.col <- "yellow"
}

# calculate overlap
ol <- calculate.overlap(list(list1,list2))

# name list 1
if(is.null(list1.n) ) {
  list1.n <- "list1"
}
# name list 2
if(is.null(list2.n) ) {
  list2.n <- "list2"
}

@@

# add symbol
res$symbol <- mcols(dds_list[[i]])$geneSymbol

# define list groups
res$isl1 <- rownames(res) %in% list1
res$isl2 <- rownames(res) %in% list2
res$isol <- rownames(res) %in% ol$a3

# order results (list ol on top)
res <- res[order(abs(res$log2FoldChange), decreasing = T),]
res <- res[order(res$isl1, decreasing = F),]
res <- res[order(res$isl2, decreasing = F),]
res <- res[order(res$isol, decreasing = F),]

@@

# name colors
keyvals <- ifelse(
  res$isol == TRUE, ol.col,
    ifelse(
  res$isl2 == TRUE, list2.col,
    ifelse(
    res$isl1 == TRUE, list1.col,
    'grey') ) )

keyvals[is.na(keyvals)] <- 'grey'
  names(keyvals)[keyvals == ol.col] <- "overlap"
  names(keyvals)[keyvals == list1.col] <- list1.n
  names(keyvals)[keyvals == list2.col] <- list2.n
  names(keyvals)[keyvals == 'grey'] <- 'other'

ev_f <- EnhancedVolcano(res,
    x = 'log2FoldChange',
    y = 'padj',
    lab = res$symbol,
    selectLab  = res$symbol[which(names(keyvals) == list2.n | names(keyvals) ==list1.n | names(keyvals) =="overlap")],
    labSize = 3,
    drawConnectors = TRUE,
    boxedLabels = FALSE,
    widthConnectors = 0.5,
    colConnectors = lcol,
    pointSize = 4.0,
    max.overlaps = 100,
    colCustom = keyvals,
    xlim = c(-xlim,xlim),
    ylim = c(0, ylim),
    ylab = "Padj (-log10)",
    title = n,
    subtitle = paste("list1:",length(list1),", list2: ",length(list2)),
    # sub = "SVF",
    # pCutoff = 10^(-60),
    # FCcutoff = 1,
    shapeCustom =shape,
    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
    legendLabels=c('Not sig.','|L2F| > 2','p-adj < 0.05',
                    'p-adj & L2F'),
    legendPosition = 'bottom',
    col=c('grey', pcol, pcol, pcol),
    legendLabSize = 8,
    legendIconSize = 2.0,
    axisLabSize = 8,
    titleLabSize = 8,
    subtitleLabSize = 8,
    captionLabSize = 8
    )
ev_f

@@

# single gene
    d <- plotCounts(dds, gene=ROC40, intgroup=c("condition","experiment","genotype","treatment","clientName"), main=s,returnData=TRUE)
    colnames(d)[1] <- "counts"

gcounts <- ggplot(d, aes(x = treatment, y = counts, fill=condition, color=condition)) +
    geom_boxplot(color="black") +
    geom_point(shape=21,color="black",aes(fill=condition),position=position_dodge(width=0.75), alpha=1) +
  # geom_text_repel(aes(label=clientName)) +
    scale_fill_manual(values=group.colors) +
    scale_color_manual(values=group.colors) +
    scale_y_continuous(trans = "log2") +
    labs(title = paste("ROC40"," in ",colData(dds)$experiment[1],sep = "")) +
 theme(axis.text.x = element_text(size = 10, face = "bold"),
        axis.text.y = element_text(color = "grey20", size = 8, angle = 0, hjust = 1, vjust = 0, face = "plain"),
        axis.title.x = element_text(color = "white", size = 12, angle = 0, hjust = .5, vjust = 0, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 12, angle = 90, hjust = .5, vjust = .5, face = "bold"))

gcounts %>% print()

@@

# ggsave(paste(pub_figures,"ROC40.pdf",sep="/"), plot = gcounts,
#   width = 5,
#   height = 4)

@@

## Make results
```{r, eval=TRUE, include=FALSE, echo=FALSE}

@@

@@

## Volcanos

### Volcanos 2.0
#### top genes
```{r top_volcano2}

i <- 1
ev1_d <- Vulcano_SK(res=res_ashr_list[[i]][5],
                    xlim=5,ylim=10,
                    list2.col = "black",list1.col = "grey35")
ev1_b <- Vulcano_SK(res=res_ashr_list[[i]][6],
                    xlim=5,ylim=10,
                    list2.col = "royalblue4",list1.col = "royalblue1")
ev1_r <- Vulcano_SK(res=res_ashr_list[[i]][7],
                    xlim=5,ylim=10,
                    list2.col = "firebrick4",list1.col = "firebrick1")

i <- 2
ev2_d <- Vulcano_SK(res=res_ashr_list[[i]][5],
                    xlim=5,ylim=50,
                    list2.col = "black",list1.col = "grey35")
ev2_b <- Vulcano_SK(res=res_ashr_list[[i]][6],
                    xlim=5,ylim=50,
                    list2.col = "royalblue4",list1.col = "royalblue1")
ev2_r <- Vulcano_SK(res=res_ashr_list[[i]][7],
                    xlim=5,ylim=50,
                    list2.col = "firebrick4",list1.col = "firebrick1")

ev1_d+ev1_b+ev1_r  +
  plot_layout(guides = "collect", axes="collect", axis_titles="collect") & 
  theme(legend.position = 'bottom', axis.title=element_text(size=8))

ev2_d+ev2_b+ev2_r  +
  plot_layout(guides = "collect", axes="collect", axis_titles="collect") & 
  theme(legend.position = 'bottom', axis.title=element_text(size=8))

# ga <- ggarrange(evb,
#                 evd+rremove("ylab"),
#                 evr+rremove("ylab"),
#                 ncol = 3, nrow = 2,common.legend = TRUE, legend = "bottom",
#                 widths = c(1.05,1,1))
# 
# 
# ggsave(ga, file="figures/Vulcano_all.pdf",
#        width = 12,
#        height = 8)

#### interaction in D-D, Bl-Bl,R-R
```{r vsvs_volcano2}

i <- 1
ev1_d <- Vulcano_SK(res=res_ashr_list[[i]][5],
                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
                    list1.n="blue", list2.n="red",
                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
                    xlim=5,ylim=10)
ev1_b <- Vulcano_SK(res=res_ashr_list[[i]][6],
                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
                    list1.n="blue", list2.n="red",
                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
                    xlim=5,ylim=10)
ev1_r <- Vulcano_SK(res=res_ashr_list[[i]][7],
                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
                    list1.n="blue", list2.n="red",
                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
                    xlim=5,ylim=10)

i <- 2
ev2_d <- Vulcano_SK(res=res_ashr_list[[i]][5],
                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
                    list1.n="blue", list2.n="red",
                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
                    xlim=5,ylim=50)
ev2_b <- Vulcano_SK(res=res_ashr_list[[i]][6],
                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
                    list1.n="blue", list2.n="red",
                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
                    xlim=5,ylim=50)
ev2_r <- Vulcano_SK(res=res_ashr_list[[i]][7],
                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
                    list1.n="blue", list2.n="red",
                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
                    xlim=5,ylim=50)

ev1_d+ev1_b+ev1_r  +
  plot_layout(guides = "collect", axes="collect", axis_titles="collect") & 
  theme(legend.position = 'bottom', axis.title=element_text(size=8))

ev2_d+ev2_b+ev2_r  +
  plot_layout(guides = "collect", axes="collect", axis_titles="collect") & 
  theme(legend.position = 'bottom', axis.title=element_text(size=8))

#### interaction in BL-D, R-D
```{r vsvs2_volcano2}

deg_list[[1]][8:9] %>% names()

i <- 1
ev1.1 <- Vulcano_SK(res=res_ashr_list[[i]][1],
                    list1=top_list[[i]][[8]], list2=top_list[[i]][[9]],
                    list1.n="blue", list2.n="red",
                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
                    xlim=5, ylim=50)
ev1.2 <- Vulcano_SK(res=res_ashr_list[[i]][2],
                    list1=top_list[[i]][[8]], list2=top_list[[i]][[9]],
                    list1.n="blue", list2.n="red",
                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
                    xlim=5,ylim=50)
ev1.3 <- Vulcano_SK(res=res_ashr_list[[i]][3],
                    list1=top_list[[i]][[8]], list2=top_list[[i]][[9]],
                    list1.n="blue", list2.n="red",
                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
                    xlim=5,ylim=50)
ev1.4 <- Vulcano_SK(res=res_ashr_list[[i]][4],
                    list1=top_list[[i]][[8]], list2=top_list[[i]][[9]],
                    list1.n="blue", list2.n="red",
                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
                    xlim=5,ylim=50)

i <- 2
ev2.1 <- Vulcano_SK(res=res_ashr_list[[i]][1],
                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
                    list1.n="blue", list2.n="red",
                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
                    xlim=5,ylim=100)
ev2.2 <- Vulcano_SK(res=res_ashr_list[[i]][2],
                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
                    list1.n="blue", list2.n="red",
                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
                    xlim=5,ylim=100)
ev2.3 <- Vulcano_SK(res=res_ashr_list[[i]][3],
                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
                    list1.n="blue", list2.n="red",
                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
                    xlim=5,ylim=100)
ev2.4 <- Vulcano_SK(res=res_ashr_list[[i]][4],
                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
                    list1.n="blue", list2.n="red",
                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
                    xlim=5,ylim=100)

ev1.1 + ev1.2 + ev1.3 + ev1.4 +
  plot_layout(guides = "collect", axes="collect", axis_titles="collect") & 
  theme(legend.position = 'bottom', axis.title=element_text(size=8))

ev2.1 + ev2.2 + ev2.3 + ev2.4 +
  plot_layout(guides = "collect", axes="collect", axis_titles="collect") & 
  theme(legend.position = 'bottom', axis.title=element_text(size=8))

ev1.1 + ev1.2 + ev2.1 + ev2.2 +
  plot_layout(guides = "collect", axes="collect", axis_titles="collect") & 
  theme(legend.position = 'bottom', axis.title=element_text(size=8))

### dark
```{r, eval=FALSE}

# dark
res <- res_ashr_list[[i]][[5]]
n <- 'WT vs. pCRY in dark'
nf <- 'figures/Volcano_dark.png'
l <- nrow(top_list[[i]][['pCRY_D.vs.WT_D']])
pcol <- "black"
lcol <- colours[1+1]

rownames(res) <- mcols(dds_list[[i]])$id.symbol
res <- res[!is.na(res$padj),]
res <- res[!is.na(res$log2FoldChange),]
